### 2013-01-05 exercise2/paper/paper.tex

%x%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage{color}
\usepackage{float}
\usepackage{placeins}
\usepackage{natbib}
% philipp: added for code listings
\usepackage{listings}
\lstset{
    captionpos=b
}
\newcommand{\Hilight}{\makebox[0pt][l]{\color{red}\rule[-4pt]{0.65\linewidth}{14pt}}}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{morefloats}
\bibliographystyle{plain}

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template
\usepackage{todonotes}

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
% \renewcommand{\thesection}{Data Set \arabic{section}}
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%   TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{ 
\normalfont \normalsize 
\textsc{Vienna University of Technology} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Experiments in Machine Learning 2 \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Sch\"{a}fer} % Your name

\date{\normalsize\today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

%----------------------------------------------------------------------------------------
%   PROBLEM 1
%----------------------------------------------------------------------------------------

\tableofcontents

%
%
%
\section{TODO}

\begin{itemize}
\item Structure for paper
\item Implementation of cutsom bn-searcher
\item documentation of cutsom bn-searcher
\item comparison and interpretation of the results
\item description of the used R and Weka features and commands
\item description of the setup/infrastructure (weka, custom searcher, R, runtime, comparison, data set, ...)
\end{itemize}

%
%
%
\section{Preface}

\paragraph{}This assignment is about extending a Bayes Network Classifier by adding a custom search algorithm. We chose the Hill Climber search algorithm for an implementation of iterative local search algorithm.

%
%
%
\section{Setup}

\paragraph{}To run the search algorithm an R script has be written. It loads Weka,builds the classifier, runs it and evaluates the model. It also shows the performance tables and the differences of the original Hill Climber implementation and our improved one.

\paragraph{} Tools needed to run the code: R (RWeka, rJava), Java (JRE).

\paragraph{} RWeka is a interface in R that wraps the Weka framework and has a really good documentation.

%
%
\section{Current Implementation}

First we take a look at the BayesNet classifier implementation. When the classifier is startet, first thing is to generate the classification model. Therefore the BayesNet classifier initializes (\texttt{initStructure()})an adjacency matrix where each attribute has an array of parents. These arrays are empty after initialization.

\paragraph{} The default SearchAlgorithm.class initializes the network like a Naive Bayes network. This shows how we can use the adjacency matrix of BayesNet:

\begin{lstlisting}
...
// initialize parent sets to have arrow from classifier node to
// each of the other nodes
for (int iAttribute = 0; iAttribute < instances.numAttributes(); iAttribute++) {
    if (iAttribute != iClass) {
        bayesNet.getParentSet(iAttribute).addParent(iClass, instances);
    }
}
...
\end{lstlisting}

\paragraph{} Where \texttt{BayesNet} is the classifier that holds the network datastructure and \texttt{getParentSet(n)} returns the \texttt{nth} row in the adjacency matrix.

%
%
%
\section{Hill Climber}

Current implementation of Weka's Hill Climber algorithm.

\begin{lstlisting}
double fScore = calcScore(bayesNet);
Operation oOperation = getOptimalOperation(bayesNet, instances);
while ((oOperation != null) && (oOperation.m_fScore > fScore)) {
	performOperation(bayesNet, instances, oOperation);
	fScore = oOperation.m_fScore;
	oOperation = getOptimalOperation(bayesNet, instances);
}  
\end{lstlisting}

\paragraph{}As shown in the listing, Hill Climber calculates an optimal solution based on the current network iteratively until no opration can be found or the new network is only slightly better than the previous one. To find the at least worse next operation all possible operations get evaluated and the best one is returned.

\section{Extended Hill Climber}

TODO: describe our extension.

\section{Comparison}

TODO:
* used datasets
* accuracy results
* runtime/performance

\bibliography{references}

\end{document}
%------------------------------------------------